### Basic HYSPLIT Trajectory Extraction  
This short vingette demonstrates how to use splitR and visualize the data spatially. See <https://github.com/rich-iannone/splitr> for more detail from the package author.

### Installation

The devtools package is required to install `splitr` from github.

```{r install}
#install.packages("devtools")  
library(devtools)  
#devtools::install_github("rich-iannone/splitr")  
library(splitr)
```

#### Basic Use

Pull trajectory files with splitR (0.4.0.9000).

```{r, eval = FALSE}
trajectory_model <- create_trajectory_model() %>%
  add_trajectory_params(
    lat = 42,
    lon = -121,
    height = 1000,
    duration = 120,
    days = seq(as.Date('2023-12-15'), as.Date('2023-12-16'), by = "day"),
    daily_hours = "12",
    direction = "backward",
    met_type = "reanalysis",
    model_height = 10000) %>%
  splitr::run_model()
```

`trajectory_model` is a list containing the trajectory data for the
given dates in December 2023. Refer here for more details on the package
and documentation: <https://github.com/rich-iannone/splitr> \
- `lat` the latitude for the HYSPLIT model initiation location

\- `lon` the longitude

\- `height` the above ground level for air parcel initiation in meters

\- `duration` the number of hours to run the HYSPLIT model backward or
forward

\- `days` the sequence of dates to model over

\- `daily_hours` the hours of the each day to initiate a trajectory
model

\- `direction` tells HYSPLIT to run forward or backward from the
beginning datetime

\- `met_type` user selection for atmospheric data downloaded from the
NOAA FTP server to force HYSPLIT (gdas1, narr, reanalysis)

\- `model_height` sets the maximum height HYSPLIT will track air parcels
in meters

The `trajectory_model` converts directly to a dataframe using
splitr::get_output_table(). Plot the trajectories via ggplot to examine
data spatially.

```{r}
library(sf)
library(maps) 
ggplot(data = basic_traj_df) +
    borders("world", xlim = c(-130, -60), ylim = c(20, 50), colour = "gray85", fill = "gray80") + # background map with the extent over North America
    geom_point(aes(x = lon, y = lat, color = traj_dt))+ # air parcel location data, colored by the datetime
    labs(title = "HYSPLIT Hourly Trajectories", subtitle = '120 hours backwards', x = "Longitude",
        y = "Latitude", color = "Datetime") +
    theme(plot.title = element_text(hjust = 0.5, size = 16),
        axis.title = element_text(size = 12))+
    theme_minimal()
```

The plot shows two trajectories initiated at 12 PM on December 15 and 16
2023. The trajectories are colored by time, lighter is more recent and
darker as the hour along trajectory approaches a maximum of 120 hours
prior to the imitation time.
